---
course: STAT 305
section: B
semester: Spring 2019
date: April 11, 2019
number: II
output:
  pdf_document:
    keep_tex: true
    template: ../../templates/exam-template.tex
    latex_engine: pdflatex
...

```{r echo=FALSE, cache=FALSE, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
knitr::opts_chunk$set(
   cache=FALSE,
   fig.width=4,
   fig.height=4,
   fig.align='center',
   out.width='.4\\linewidth',
   out.height='.4\\linewidth',
   echo=FALSE
)
theme_set(theme_bw())
```

```{r include=FALSE}
```

\question[2] For continuous distributions, the probability density function is \fillin\ continuous and the cumulative density function is \fillin\ continuous.
\begin{oneparchoices}
\choice \textbf{always} and \textbf{always}
\choice \textbf{always} and \textbf{sometimes}
\choice \textbf{sometimes} and \textbf{always}
\choice \textbf{sometimes} and \textbf{sometimes}
\end{oneparchoices}
\vspace{1cm}

\question[3]
Circle the name of the distribution which best matches the plot of the **probability density function**:

```{r}
ggplot() +
   geom_path(aes(x=c(-3, -2), y=c(0, 0))) +
   geom_path(aes(x=c(2, 3), y=c(0, 0))) +
   stat_function(data=data.frame(x=0), mapping=aes(x=x), fun=dunif, args=list(min=-2, max=2), xlim=c(-2, 2)) +
   ylab('f(x)') +
   xlab('x')
```
\begin{oneparchoices}
\choice standard normal
\choice normal 
\choice exponential 
\choice poisson
\choice uniform
\end{oneparchoices}

\question[3]
Circle the name of the distribution which best matches the plot of the **cumulative density function**:

```{r}
ggplot() + 
   stat_function(data=data.frame(x=0), mapping=aes(x=x), fun=pnorm, args=list(mean=2, sd=1)) +
   xlim(-1, 5) + 
   ylab('F(x)') +
   xlab('x')
```
\begin{oneparchoices}
\choice standard normal
\choice normal 
\choice exponential 
\choice binomial
\choice uniform
\end{oneparchoices}
\vspace{1cm}


\newpage
\question[4] The left plot depicts the **cumulative density function** of a random variable $X$. In the plot on the right, sketch the corresponding probability density function.
```{r, fig.width=9, out.width='0.9\\linewidth'}
p0 <- ggplot() + 
   geom_path(aes(x=c(-1, 0, 1, 2, 3, 4), y=c(0, 0, 0.25, 0.25, 1, 1))) +
   ylab('cumulative density function, F(x)') +
   xlab('x')
p1 <- ggplot(data=NULL, aes(x=c(-1, 4), y=c(0, 1))) + 
   theme_bw() + xlab('x') + ylab('probability density function, f(x)') +
   geom_blank()
grid.arrange(p0, p1, ncol=2)
```

\vspace{1cm}

\question Suppose $X$ is a discrete random variable with following probability function:
$$f(x) = \begin{cases} 0.1 |x| + 0.08 & x = -2, -1, 0, 1, 2 \\ 0 & o.w. \end{cases}$$

\qparts{
\part[2] Find $E(X)$ \vspace{2.5cm}
\part[2] Find $P(|X| = 2)$ \vspace{2.5cm}
\part[2] Find $P(X \ge 0)$ \vspace{2.5cm}
}

\newpage

\question 
Let $X$ be a binomial random variable with $n = 6$ and $p = 0.2$. 
\qparts{
\part[2] Find $P(X = 3)$ \vspace{2cm}
\part[2] Find $P(X > 4)$ \vspace{2cm}
\part[2] Find $E(X)$ \vspace{2cm}
\part[2] Find $Var(X)$ \vspace{2cm}
}

\question 
Suppose $Y$ is an exponential random variable with $E(Y) = 2$.
\qparts{
\part[2] Find $P(Y \le 3)$ \vspace{2cm}
\part[2] Find $P(1 \le Y \le 3)$ \vspace{3cm}
\part[2] What is the cumulative density function of $Y$ \vspace{3cm}
\part[2] Find $Var(Y)$ \vspace{2cm}
}

\newpage

\question

Let $X$ be a normal random variable with a mean of 50 and a varaince of 25 (i.e., $X \sim N(50, 25)$) and let $Z$ be a random variable following a standard normal distribution. Find the following probabilities (note: the Standard Normal Probabilites table attached at to the exam will helpful): \vspace{1cm}
\qparts{
  \part[2] $P(Z > 0.1)$ \vspace{2cm}
  \part[2] $P(|Z| \le 1.25)$ \vspace{2cm}
  \part[2] $P(1 < Z < 3)$ \vspace{2cm}
  \part[2] $P(X < 40)$ \vspace{3cm}
  \part[2] $P(|X-50| \le 12.5)$ \vspace{3cm}
  \part[5] Find the value $a$ so that $P(X > a) = 0.85$
}

\newpage

\question 
We are going to play a game involving four bags. The first bag contains 10 lettered wooden tiles (3 letterd A, 2 lettered B, and 5 lettered C). 
The other bags are labelled \textbf{Bag A}, \textbf{Bag B}, \textbf{Bag C} and contain the following:

\qitems{
\item \textbf{Bag A}: 25 colored marbles: 16 Blue, 4 Red, 3 Green, and 2 Yellow.
\item \textbf{Bag B}: 50 colored marbles: 10 Blue, 38 Red, 1 Green, and 1 Yellow.
\item \textbf{Bag C}: 30 colored marbles: 1 Blue, 11 Red, 17 Green, and 1 Yellow.
}

Once you have learned the contents of the bags, we play the following game:

\qitems{
\item First, I randomly select a tile from the bag of tiles (you do not see the tile). 
\item Next, I take the bag of marbles with the label matching the letter on my tile (you do not see which bag I take).
\item Then, I randomly select a marble from the bag I took.
\item Finally, I show you the marble. 
}

You win the game by correctly guessing the letter on the tile I drew. \vspace{.3cm}

\qparts{
\part[2] What is the probability that I start the game by drawing a tile with the letter A? \vspace{3cm}
\part[2] What is the probability that I draw a blue marble \textit{given} that I drew a tile with the letter A? \vspace{3cm}
\part[2] What is the probability that I start the game by drawing a tile with the letter A and then draw a blue marble? \vspace{3cm} 
\part[2] What is the probability that I start the game by drawing a tile with the letter C and then draw a blue marble? \vspace{3cm} 
\part[2] What is the probability that the marble I show you a blue marble? \vspace{3cm}
\part[2] What is the probability that I drew a tile with the letter A \textit{given} that I show you a blue marble? \vspace{3cm}
\part[2] What is the probability that I drew a tile with the letter B \textit{given} that I show you a green marble? \vspace{3cm}
}

\newpage

\question 

Suppose that $X$ is a continuous random variable with probability density function (pdf):
$$ f(x) = \begin{cases} c(4 - x^2) &  -2 \le x \le 2 \\ 0 & \text{otherwise} \end{cases} $$
where $c$ is a constant (not necessarily positive).

\qparts{
\part[2] What is the value of $c$ if $f(x)$ is a valid probability density function? \vspace{3cm}
\part[5] Sketch the probability density function using the grid below (including the points $(-2, f(2))$, $(0, f(0))$, and $(2, f(2))$).

```{r, fig.width=8, fig.height=3, out.width='0.8\\linewidth', out.height='0.3\\linewidth'}
ggplot(data=NULL, aes(x=c(-2.5, 2.5), y=c(0, 5))) + 
   theme_bw() + xlab('x') + ylab('f(x)') +
   geom_blank()
```
\part[4] What is the cumulative density function, $F(x)$? \vspace{3.5cm}
\part[2] What is the probability that $X$ takes a value greater than 0.5? \vspace{2cm}
\part[2] What is the probability that $X$ takes a value between 0.5 and 1? \vspace{2cm}

}
