---
title: Competency Quiz 4
course: Stat 305
semester: Spring 2019
due-date: Tuesday, April 18 # Like January 4
output:
  pdf_document:
    keep_tex: true
    template: ../../templates/cq-template.tex
    latex_engine: pdflatex
...
```{r echo=FALSE, cache=FALSE, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
knitr::opts_chunk$set(
   cache=FALSE,
   fig.width=4.5,
   fig.height=2,
   fig.align='center',
   out.width='.9\\linewidth',
   out.height='.4\\linewidth',
   echo=FALSE
)
```

1. **Competency Topic: Discrete Random Variables**

   Suppose that $X$ is a discrete random variable with the following probability function:
   $$f(x) = \begin{cases} \dfrac{x^2}{c} & x = -3, -2, -1, 0, 1, 2, 3 \\ 0 & o.w \end{cases}$$
   where $c$ is a constant.
   a. Find the value of $c$ that makes $f(x)$ a valid probability function. \vspace{4cm}
   b. Find $P(X \ge 2)$. \vspace{4cm}
   c. Find $E(X)$.

\newpage

2. **Competency Topic: Continuous Random Variables**

   The mean-0 Laplace distribution is continuous distribution with the following probability density function:
   $$f(x) = \dfrac{1}{2\alpha} \exp\left(-\dfrac{|x|}{\alpha}\right) \; -\infty < x < \infty$$
   where $\alpha$ is a parameter which takes positive values (note: $|x|$ is the absolute value of $x$).
   a. Show that regardless of the value of $\alpha$ the pdf above is symmetric (that is, show that $f(-x) = f(x)$). \vspace{3cm}
   b. Using the plot below, provide a rough sketch of the following pdfs: 
   - (1) the pdf of a variable with $\alpha = 0.5$.
   - (2) the pdf of a variable with $\alpha = 1$.
   (for the sketch, show the values of the pdfs when $x = -2, -1, 0, 1, 2$)
```{r}
ggplot() + geom_blank(data=NULL, aes(x=c(-2,2), y=c(0,0.5))) + xlab('x') + ylab('f(x)') + theme_bw()
```
   c. Based on your sketch, would a random variable with $\alpha = 0.5$ have a larger or smaller variance than a random variable with $\alpha = 1$?

\newpage

3. **Competency Topic: Joint Distributions**

   Suppose that $X$ is a random variable with an exponential distribution with mean $\lambda$. That is
   $$f_X(x) = \begin{cases} \dfrac{1}{\lambda} \exp\left(-\dfrac{x}{\lambda}\right) & x \ge 0 \\ 0 & otherwise \end{cases}$$
   and suppose that a random variable $Y$ follows an exponential distribution that depends on the value taken by $X$ so that
   $$f_Y|X(y|x) = \begin{cases} \frac{1}{x} & 0 \le y \le x \\ 0 & otherwise \end{cases}$$
   a. Sketch the conditional probability density function of $Y$ given that $X = 4$.
```{r}
ggplot() + geom_blank(data=NULL, aes(x=c(-2,5), y=c(0,1))) + xlab('y') + ylab('f(y)') + theme_bw() + theme(axis.text.x=element_blank())
```
   b. Find the joint probability function, $f_{XY}(x, y)$. \vspace{3cm}

\newpage

4. **Competency Topic: Functions of Random Variables**

   Suppose that $W_1, W_2, ...$ are independent random variables each with the same expected value $\mu$ and variance $\sigma^2$.
   We define a random variable $U_n$ as a linear combination of $n$ of these random variables using:
   $$U_n = 2^{-1} W_1 + 2^{-2} W_2 + 2^{-3} W_3 + \ldots + 2^{-n} W_n$$
   note: if $r \ge 1$, the $r^{-1} + r^{-2} + \ldots r^{-n} = \sum_{k=1}^n r^{-k} = \dfrac{r^{n-1} - 1}{r^n - r^{n-1}}$
   a. Find an expression for $E(U_n)$ (hint: it will contain $\mu$ and $n$). \vspace{5cm}
   b. Find an expression for $Var(U_n)$ (hint: it will contain $\sigma^2$ and $n$).
