---
course: STAT 305
section: B
semester: Spring 2019
date: April 11, 2019
number: II
output:
  pdf_document:
    keep_tex: true
    template: ../../templates/exam-template.tex
    latex_engine: pdflatex
...

```{r echo=FALSE, cache=FALSE, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
knitr::opts_chunk$set(
   cache=FALSE,
   fig.width=4,
   fig.height=4,
   fig.align='center',
   out.width='.4\\linewidth',
   out.height='.4\\linewidth',
   echo=FALSE
)
```

\question[2] For discrete distributions, the probability function is \fillin\ and the cumulative probability function is \fillin.
\begin{oneparchoices}
\choice discontinuous and discontinuous
\choice discontinuous and continuous
\choice continuous and continuous
\choice continuous and discontinuous
\end{oneparchoices}
\vspace{1cm}

\question[3]
Circle the name of the distribution which best matches the plot of the probability density function:

```{r}
x <- seq(-3, 3, by=0.1)
ggplot() + geom_path(aes(x=x, y=dnorm(x, 0, 1))) + theme_bw() + xlab('x') + ylab('f(x)')
```
\begin{oneparchoices}
\choice exponential 
\choice normal 
\choice uniform
\choice binomial
\end{oneparchoices}
\vspace{1cm}

\question[3]
Circle the name of the distribution which best matches the plot of the cumulative density function:

```{r}
x <- seq(-3, 3, by=0.1)
ggplot() + 
   geom_path(aes(x=c(x[x<0], 0), y=c(dexp(x[x<0], 1), 0))) + 
   geom_path(aes(x=x[x>=0], y=dexp(x[x>=0], 1))) + 
   theme_bw() + xlab('x') + ylab('F(x)')
```
\begin{oneparchoices}
\choice exponential 
\choice normal 
\choice uniform
\choice binomial
\end{oneparchoices}

\newpage
\question[4] The left plot depicts the probability density function of a step-uniform random variable. In the plot on the right, sketch the corresponding cumulative density function.
```{r, fig.width=9, out.width='0.9\\linewidth'}
x <- seq(-3, 3, by=0.1)
p0 <- ggplot() + 
   geom_path(aes(x=c(-3, -2), y=c(0, 0))) + 
   geom_path(aes(x=c(-2, 0), y=c(.2, .2))) + 
   geom_path(aes(x=c(0, 2), y=c(.3, .3))) + 
   geom_path(aes(x=c(2, 3), y=c(0, 0))) + 
   theme_bw() + xlab('x') + ylab('f(x)')
p1 <- ggplot(data=NULL, aes(x=c(-3, 3), y=c(0, 1))) + 
   theme_bw() + xlab('x') + ylab('F(x)') +
   geom_blank()
grid.arrange(p0, p1, ncol=2)
```

\vspace{1cm}

\question Suppose $X$ is a discrete random variable with following probability function:
$$ f(x) = \begin{cases} 0.1 & x = -2, 0, 2 \\ 0.35 & x = -1, 1 \\ 0 & o.w. \end{cases} $$

\qparts{
\part[2] Find $P(X = 0)$ \vspace{2.5cm}
\part[2] Find $P(X \le 0)$ \vspace{2.5cm}
\part[2] Find $P(X \ne 0)$ \vspace{2.5cm}
}

\newpage

\question 
Two manufacturers (A and B) of the same electrical circuit board are being used by a company. 
The boards have five circuit connectors which can either be defective or non-defective (meaning that each circuit board could have 0, 1, 2, 3, 4, or 5 nondefective connectors).
The company knows that the manufacturer matters when it comes to the chance that the connectors are defective.
Unfortunately, all the circuit boards have been placed in the same container and there is no clear way to tell them apart.
We do know the following though:
$$P(\text{``There are }k\text{ defective connectors on the board"} | \text{``The board is manufacturer A"}) = \frac{5!}{(5-k)! k!} \left(\frac{1}{4}\right)^k \left(\frac{3}{4}\right)^{5 - k} $$  
$$P(\text{``There are }k\text{ defective connectors on the board"} | \text{``The board is manufacturer B"}) = \frac{5!}{(5-k)! k!} \left(\frac{1}{2}\right)^k \left(\frac{1}{2}\right)^{5 - k} $$  
Suppose that we also know that there were 300 boards of manufacturer A and 200 boards of manufacturer B.

\qparts{
\part[2] Find the probability that a randomly chosen board is from manufacturer A. \vspace{3cm}
\part[2] Find the probability that a randomly chosen board has no defective connectors \textit{given} that it is from manufacturer A. \vspace{3cm}
\part[2] Find the probability that a randomly chosen board is from manufacturer A and has 0 defective connectors. \vspace{3cm}
\part[2] Find the probability that a randomly chosen board is from manufacturer B and has 0 defective connectors. \vspace{3cm}
\part[2] Find the probability that a randomly chosen board has 0 defective connectors. \vspace{3cm}
\part[2] Find the probability that a randomly chosen board is from manufacturer A \textit{given} that it has 0 defective connectors. \vspace{5cm}
\part[2] Find the probability that a randomly chosen board is from manufacturer B \textit{given} that it has 2 defective connectors. \vspace{5cm}
}

\newpage

\question 
Let $X$ be a poisson random variable with $\lambda = 6$. 
\qparts{
\part[2] Find $P(X = 3)$ \vspace{2cm}
\part[2] Find $P(X < 2)$ \vspace{2cm}
\part[2] Find $E(X)$ \vspace{2cm}
\part[2] Find $Var(X)$ \vspace{2cm}
}

\question 
Let $Y$ be an exponential random variable with mean 2.
\qparts{
\part[2] Find $P(Y \le 3)$ \vspace{2cm}
\part[2] Find $P(1 \le Y \le 3)$ \vspace{3cm}
\part[2] Find $E(Y)$ \vspace{2cm}
\part[2] Find $Var(Y)$ \vspace{2cm}
}

\newpage

\question

Let $X$ be a normal random variable with a mean of 2 and a varaince of 9 (i.e., $X \sim N(2,9)$) and let $Z$ be a random variable following a standard normal distribution. Find the following probabilities (note: Table B-3 will be helpful): \vspace{1cm}
\qparts{
  \part[2] $P(Z \le 2)$ \vspace{2cm}
  \part[2] $P(|Z| \ge 1)$ \vspace{2cm}
  \part[2] $P(0 \le Z < 3)$ \vspace{2cm}
  \part[2] $P(X < 3)$ \vspace{3cm}
  \part[2] $P(|X-2| \le 2.5)$ \vspace{3cm}
  \part[5] Find the value $a$ so that $P(X < a) = .95$ (hint: start with a standard normal)
}

\newpage

\question 

Suppose that $X$ is a continuous random variable with probability density function (pdf):
$$ f(x) = \begin{cases} 0.25 + c x^2 &  -1 < x < 1 \\ 0 & \text{otherwise} \end{cases} $$
where $c$ is a constant (not necessarily positive).

\qparts{
\part[2] What is the value of $c$ if $f(x)$ is a valid probability density function? \vspace{2cm}
\part[5] Sketch the probability density function using the grid below (including the points $(-1, f(1))$, $(0, f(0))$, and $(1, f(1))$).

```{r, fig.width=9, out.width='0.9\\linewidth'}
ggplot(data=NULL, aes(x=c(-1, 1.5), y=c(1.5, 1.5))) + 
   theme_bw() + xlab('x') + ylab('f(x)') +
   geom_blank()
```
\part[4] What is the cumulative density function, $F(x)$? \vspace{3cm}
\part[2] What is the probability that $X$ takes a value greater than 1? \vspace{2cm}
\part[2] What is the probability that $X$ takes a value between 0 and 1? \vspace{2cm}
}
